---
title: "CPM checks"
output:
  html_document:
    toc: no
    toc_depth: 3
    number_sections: false
    code_folding: hide
    theme: paper
---

<!--
Comparisons between different versions of Bayes CPM code and 
between rstanarm and hardcoded CPM
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
libs <- c("rstan", "rstanarm", "rms", "dplyr", "stringr", "readr", "bayesplot","ordinal","MASS")
invisible(lapply(libs, library, character.only = TRUE))

# repro & update these functions for general Stan output
#dir<-getwd()
#source(file.path(dir,"rstanarm_ord_functions.r"))

set.seed(24834)

# call this once to distribute MCMC chains across cpu cores:
options(mc.cores=parallel::detectCores())
```

Read in CPM models

```{r}
# compile ordinal models

# concentration (alpha) is unspecified, default is ???
if (0){
ord_mod_file0<-read_file(file.path(getwd(),"ordinal_model_0.stan"))
ord_mod0 <- stan_model(model_code = ord_mod_file0)
saveRDS(ord_mod0, file = file.path(getwd(),"ordinal_model_0.rds"))
}

ord_mod0<-readRDS(file.path(getwd(),"ordinal_model_0.rds"))

# concentration (alpha) is given as a scalar parameter along with in data
if (0){
ord_mod_file1<-read_file(file.path(getwd(),"ordinal_model_1.stan"))
ord_mod1 <- stan_model(model_code = ord_mod_file1)
saveRDS(ord_mod1, file = file.path(getwd(),"ordinal_model_1.rds"))
}

ord_mod1 <-readRDS(file.path(getwd(),"ordinal_model_1.rds"))

# concentration (alpha) is estimated with gamma(2,2) prior or exp(1) prior
if (0){
ord_mod_file2<-read_file(file.path(getwd(),"ordinal_model_2.stan"))
ord_mod2 <- stan_model(model_code = ord_mod_file2)
saveRDS(ord_mod2, file = file.path(getwd(),"ordinal_model_2.rds"))

# ord_mod_file2b<-read_file(file.path(getwd(),"ordinal_model_2b.stan"))
# ord_mod2b <- stan_model(model_code = ord_mod_file2b)
# saveRDS(ord_mod2b, file = file.path(getwd(),"ordinal_model_2b.rds"))
# 
# ord_mod_file2c<-read_file(file.path(getwd(),"ordinal_model_2c.stan"))
# ord_mod2c <- stan_model(model_code = ord_mod_file2c)
# saveRDS(ord_mod2c, file = file.path(getwd(),"ordinal_model_2c.rds"))
}

ord_mod2 <-readRDS(file.path(getwd(),"ordinal_model_2.rds"))
# ord_mod2b <-readRDS(file.path(getwd(),"ordinal_model_2b.rds"))
# ord_mod2c <-readRDS(file.path(getwd(),"ordinal_model_2c.rds"))

```

## example with no covariates

```{r ex0, cache=TRUE}
set.seed(473)
n <- 100
y <- round(runif(n), 4)
x <- rep(1,n)

# number unique values; no ties
n_uniq<-length(unique(y))
cuts <- n_uniq - 1


dat0<-data.frame(y,x)

# using lrm
# NOTE: f_lrm_alt <- lrm(y ~ x, eps=1e-5) doesn't work with x as constant
f_lrm <- lrm(y ~ 1, eps=1e-5)

y2 <- y-100
f_lrm2 <- lrm(y2 ~ 1, eps=1e-5)

y3 <- log(y)
f_lrm3 <- lrm(y3 ~ 1, eps=1e-5)

# intercepts are the same regardless of shift and transformation



# using cumulative link model (clm) from ordinal package
f_clm <- clm(factor(y) ~ 1, data=dat0) 
#!! f_clm_alt <- clm(factor(y) ~ x, data=dat0)

#!! all.equal(summary(f_clm)$alpha, summary(f_clm_alt)$alpha)

cbind(-coef(f_lrm),-coef(f_lrm2), -coef(f_lrm3), summary(f_clm)$alpha)

# using orm doesn't work w/ no covars
# f_orm <- orm(y ~ 1, eps=1e-5) # doesn't work with only 'intercept'

#f_orm_alt <- orm(y ~ x, eps=1e-5)

#cbind(-coef(f_orm_alt)[1:cuts], -coef(f_lrm),
#      summary(f_clm)$alpha, summary(f_clm_alt)$alpha)

# doesn't work with no covariates
# b <- stan_polr(ordered(y) ~ 1, data = dat0,
#               prior=NULL, prior_counts = dirichlet(1), method="logistic",
#               adapt_delta = 0.99)

# only expect Bayes model MAP to match MLE for intercepts if use noninformative prior 
# for intercepts and betas AND number of observations in each cell is moderate 
# see Johnson & Albert, Chp 4, pg 132
mod_data0 <- list(N=nrow(dat0),
                 ncat=length(unique(dat0$y)),
                 Ylev=as.numeric(ordered(dat0$y)),
                 link=1,
                 K=ncol(dat0[,c("x"),drop=FALSE]),
                 Q=dat0[,c("x"),drop=FALSE],
                 alpha=1)

f_stan0 <- optimizing(ord_mod0, data=mod_data0, seed=12345)

f_stan0$par[paste0("cutpoints[",1:cuts,"]")]

f_stan1 <- sampling(ord_mod1, data=mod_data0, seed=12345, 
                 iter=3000, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

f_stan1_mn <- summary(f_stan1, pars=paste0("cutpoints[",1:cuts,"]"), probs=c(0.5))$summary[,'mean']

f_stan1_md <-summary(f_stan1, pars=paste0("cutpoints[",1:cuts,"]"), probs=c(0.5))$summary[,'50%']
```

```{r, eval=FALSE, echo=FALSE}
#print(f_stan1, pars=paste0("cutpoints[",1:cuts,"]"), probs=c(0.5))

plot(f_stan1, pars=paste0("cutpoints[",1:25,"]"))

# mean
#plot(-coef(f_lrm),1:99)
#points(f_stan1_mn,1:99, col="red")

# difference intercept MLEs vs. intercept post. mean
plot(1:99, -coef(f_lrm)-f_stan1_mn)

# median
#plot(-coef(f_lrm),1:99)
#points(f_stan1_md,1:99, col="green")

# ecdf?
plot(-coef(f_lrm),sort(y)[1:99])

# difference alpha MLEs vs. alpha post. median
plot(1:99, -coef(f_lrm)-f_stan1_md)

# specify larger concentration (alpha)
mod_data1 <- list(N=nrow(dat0),
                 ncat=length(unique(dat0$y)),
                 Ylev=as.numeric(ordered(dat0$y)),
                 link=1,
                 K=ncol(dat0[,c("x"),drop=FALSE]),
                 Q=dat0[,c("x"),drop=FALSE],
                 alpha=5)

f_stan_alp5 <- sampling(ord_mod1, data=mod_data1, seed=12345, 
                 iter=3000, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

f_stan_alp5_mn <- summary(f_stan_alp5, pars=paste0("cutpoints[",1:cuts,"]"))$summary[,'mean']

plot(-coef(f_lrm),1:99)
points(f_stan_alp5_mn,1:99, col="red")

# difference intercept MLEs vs. intercept post. mean
plot(1:99, -coef(f_lrm)-f_stan_alp5_mn)

# specify smaller concentration (alpha) 

#!! alpha = 0 is like uninformative (flat) prior, right?
#!! make model with unassigned prior for alpha  --> might not have proper posterior
mod_data2 <- list(N=nrow(dat0),
                 ncat=length(unique(dat0$y)),
                 Ylev=as.numeric(ordered(dat0$y)),
                 link=1,
                 K=ncol(dat0[,c("x"),drop=FALSE]),
                 Q=dat0[,c("x"),drop=FALSE],
                 alpha=0.1)

f_stan_alp01 <- sampling(ord_mod1, data=mod_data2, seed=12345, 
                 iter=3000, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

f_stan_alp01_mn <- summary(f_stan_alp01, pars=paste0("cutpoints[",1:cuts,"]"))$summary[,'mean']

plot(-coef(f_lrm),1:99)
points(f_stan_alp01_mn,1:99, col="red")

# difference intercept MLEs vs. intercept post. mean
plot(1:99, -coef(f_lrm)-f_stan_alp01_mn)


# estimate alpha
f_stan2 <- sampling(ord_mod2, data=mod_data0, seed=12345, 
                 iter=3000, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

f_stan2_mn <- summary(f_stan2, pars=paste0("cutpoints[",1:cuts,"]"))$summary[,'mean']

plot(-coef(f_lrm),1:99)
points(f_stan2_mn,1:99, col="red")

# difference intercept MLEs vs. intercept post. mean
plot(1:99, -coef(f_lrm)-f_stan2_mn)

```



## orm() example 1

```{r ex1a, cache=TRUE}
set.seed(3758)
n <- 100
y <- round(runif(n), 2)
x1 <- sample(c(-1,0,1), n, TRUE)
x2 <- sample(c(-1,0,1), n, TRUE)

dat_a <- data.frame(y=ordered(y),x1,x2)

# orm
fit_orm_a <- orm(y ~ x1 + x2, eps=1e-5)


## Compare stan_polr to hardcoded stan model 1 (fixed alpha) and 2 (estimate alpha)

# stan_polr
fit_polr_a <- stan_polr(y ~ x1 + x2, data = dat_a, seed=23145, 
                       iter=3500, warmup=1500, chains=2,
                       prior=NULL, prior_counts = dirichlet(1), method="logistic",
                       adapt_delta = 0.8)

# hardcoded stan model data
mod_data_a <- list(N=nrow(dat_a),
                 ncat=length(unique(dat_a$y)),
                 Ylev=as.numeric(dat_a$y),
                 link=1,
                 K=ncol(dat_a[,c("x1","x2")]),
                 Q=dat_a[,c("x1","x2")],
                 alpha=1)

fit_ord0_a <- sampling(ord_mod0, data=mod_data_a, seed=23145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_ord1_a <- sampling(ord_mod1, data=mod_data_a, seed=23145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_ord2_a <- sampling(ord_mod2, data=mod_data_a, seed=23145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

# check trace plots
mcmc_trace(fit_polr_a, pars=c("x1","x2"))
traceplot(fit_ord0_a, pars=c("b[1]","b[2]","alpha"))
traceplot(fit_ord1_a, pars=c("b[1]","b[2]"))
traceplot(fit_ord2_a, pars=c("b[1]","b[2]","alpha"))

# check summaries
fit_orm_a 
summary(fit_polr_a, pars=c("x1","x2"), digits=4)
summary(fit_ord0_a,pars=c("b[1]","b[2]","alpha"))$summary
summary(fit_ord1_a,pars=c("b[1]","b[2]"))$summary
summary(fit_ord2_a,pars=c("b[1]","b[2]","alpha"))$summary

#! stan_polr, ord_mod1 and ord_mod2 are somewhat similar (and also match orm) w/ dirchlet(1) and alpha=1 and estmating alpha, respectively
#! ord_mod2 is a bit off and alpha est is higher than 1 (approx 4)
#! is gamma prior too strong??
# hist(extract(fit_ord2_a,pars="alpha")$alpha)
```

Add table comparing 5 models (orm, stan_polr w/ dirichlet(1), ord_mod0 w noninformative (unspecified) alpha, ord_mod1 with alpha=1, ord_mod2 with gamma(2,2) prior on alpha


```{r ex1b}
# compare alpha=1/2, alpha=1/ylevs, alpha=2

if(0){
#Note: stan_polr and ord_mod1 don't work w/ improper alpha=0 concentration param
fit_polr_b0 <- stan_polr(y ~ x1 + x2, data = dat_a, seed=3145, iter=3500,
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(0), method="logistic",
               adapt_delta = 0.9)

mod_data_b0 <- list(N=nrow(dat_a),
                 ncat=length(unique(dat_a$y)),
                 Ylev=as.numeric(dat_a$y),
                 link=1,
                 K=ncol(dat_a[,c("x1","x2")]),
                 Q=dat_a[,c("x1","x2")],
                 alpha=0)

fit_ord1_b0 <- sampling(ord_mod1, data=mod_data_b0, seed=3145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.9))
}


fit_polr_b <- stan_polr(y ~ x1 + x2, data = dat_a, seed=3145, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(1/2), method="logistic",
               adapt_delta = 0.9)

ylevs<-length(unique(dat_a$y)) 
fit_polr_b2 <- stan_polr(y ~ x1 + x2, data = dat_a, seed=3145, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(1/ylevs), method="logistic",
               adapt_delta = 0.9)


fit_polr_c <- stan_polr(y ~ x1 + x2, data = dat_a, seed=1745, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(2), method="logistic",
               adapt_delta = 0.9)

mod_data_b <- list(N=nrow(dat_a),
                 ncat=length(unique(dat_a$y)),
                 Ylev=as.numeric(dat_a$y),
                 link=1,
                 K=ncol(dat_a[,c("x1","x2")]),
                 Q=dat_a[,c("x1","x2")],
                 alpha=1/2)

mod_data_b2 <- list(N=nrow(dat_a),
                 ncat=length(unique(dat_a$y)),
                 Ylev=as.numeric(dat_a$y),
                 link=1,
                 K=ncol(dat_a[,c("x1","x2")]),
                 Q=dat_a[,c("x1","x2")],
                 alpha=1/ylevs)

mod_data_c <- list(N=nrow(dat_a),
                 ncat=length(unique(dat_a$y)),
                 Ylev=as.numeric(dat_a$y),
                 link=1,
                 K=ncol(dat_a[,c("x1","x2")]),
                 Q=dat_a[,c("x1","x2")],
                 alpha=2)

fit_ord1_b <- sampling(ord_mod1, data=mod_data_b, seed=3145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.9))

fit_ord1_b2 <- sampling(ord_mod1, data=mod_data_b2, seed=3145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.9))

fit_ord1_c <- sampling(ord_mod1, data=mod_data_c, seed=1745, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.9))

# using dirichlet(1/2) or alpha=1/2
summary(fit_polr_b, pars=c("x1","x2"), digits=4)
summary(fit_ord1_b,pars=c("b[1]","b[2]"))$summary

plot(fit_polr_b, pars=c("x1","x2"), prob=0.8, prob_outer=0.95)
plot(fit_ord1_b,pars=c("b[1]","b[2]"), ci_level = 0.8, outer_level=0.95)

# using dirichlet(1/ylevs) or alpha=1/ylevs
summary(fit_polr_b2, pars=c("x1","x2"), digits=4)
summary(fit_ord1_b2,pars=c("b[1]","b[2]"))$summary

plot(fit_polr_b2, pars=c("x1","x2"), prob=0.8, prob_outer=0.95)
plot(fit_ord1_b2,pars=c("b[1]","b[2]"), ci_level = 0.8, outer_level=0.95)

# using dirichlet(2) or alpha=2
summary(fit_polr_c, pars=c("x1","x2"), digits=4)
summary(fit_ord1_c,pars=c("b[1]","b[2]"))$summary

plot(fit_polr_c, pars=c("x1","x2"), prob=0.8, prob_outer=0.95)
plot(fit_ord1_c,pars=c("b[1]","b[2]"), ci_level = 0.8, outer_level=0.95)
```

```{r}
# check match between stan_polr and hardcoded stan model
#!! need to figure out why stan_polr with dirichlet(alpha=1/2) doesn't match ord_mod1 with alpha=1/2
fit_orm_a
summary(fit_polr_a, pars=c("x1","x2"), digits=4) # dirichlet(1)
summary(fit_polr_b, pars=c("x1","x2"), digits=4) # dirichlet(1/2)
summary(fit_polr_b2, pars=c("x1","x2"), digits=4) # dirichlet(1/ylevs)
summary(fit_polr_c, pars=c("x1","x2"), digits=4) # dirichlet(2)

fit_orm_a
summary(fit_ord1_a, pars=c("b[1]","b[2]"))$summary # alpha=1 <-- somewhat close to orm
summary(fit_ord1_b, pars=c("b[1]","b[2]"))$summary # alpha=1/2 <-- somewhat close to orm
summary(fit_ord1_b2, pars=c("b[1]","b[2]"))$summary # alpha=1/ylevs <-- very close to orm!!
summary(fit_ord1_c, pars=c("b[1]","b[2]"))$summary # alpha=2 <-- somewhat close to orm

summary(fit_ord2_a, pars=c("b[1]","b[2]","alpha"))$summary # est alpha
```

```{r}
## compare models using outcome with only 4 categories
set.seed(1342)
n <- 50
y <- round(runif(n), 1)
x1 <- sample(c(-1,0,1), n, TRUE)
x2 <- sample(c(-1,0,1), n, TRUE)

dat_b <- data.frame(y=ordered(y),x1,x2)


# orm
fit_orm_bb <- orm(y ~ x1 + x2, eps=1e-5)


## Compare stan_polr to hardcoded stan model 1 (fixed alpha) and 2 (estimate alpha)

# stan_polr
fit_polr_bb <- stan_polr(y ~ x1 + x2, data = dat_b, seed=23145, 
                       iter=3500, warmup=1500, chains=2,
                       prior=NULL, prior_counts = dirichlet(1), method="logistic",
                       adapt_delta = 0.8)


ylevs<-length(unique(dat_b$y)) 
fit_polr_bb2 <- stan_polr(y ~ x1 + x2, data = dat_b, seed=3145, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(1/ylevs), method="logistic",
               adapt_delta = 0.9)

# hardcoded stan model data
mod_data_bb <- list(N=nrow(dat_b),
                 ncat=length(unique(dat_b$y)),
                 Ylev=as.numeric(dat_b$y),
                 link=1,
                 K=ncol(dat_b[,c("x1","x2")]),
                 Q=dat_b[,c("x1","x2")],
                 alpha=1)


mod_data_bb2 <- list(N=nrow(dat_b),
                 ncat=length(unique(dat_b$y)),
                 Ylev=as.numeric(dat_b$y),
                 link=1,
                 K=ncol(dat_b[,c("x1","x2")]),
                 Q=dat_b[,c("x1","x2")],
                 alpha=1/ylevs)

fit_ord1_bb <- sampling(ord_mod1, data=mod_data_bb, seed=23145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_ord1_bb2 <- sampling(ord_mod1, data=mod_data_bb2, seed=23145, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_orm_bb
summary(fit_polr_bb, pars=c("x1","x2"), digits=4) # dirichlet(1)
summary(fit_polr_bb2, pars=c("x1","x2"), digits=4) # dirichlet(1/ylevs)
summary(fit_ord1_bb, pars=c("b[1]","b[2]"))$summary # alpha=1
summary(fit_ord1_bb2, pars=c("b[1]","b[2]"))$summary # alpha=1/ylevs

```

```{r, eval=FALSE}
# check Dirichlet density?
library(plotly)
library(MCMCpack)

rdirichlet(10,c(1,1,1,1,1))
rdirichlet(10,c(10,10,10,10,10))
rdirichlet(10,c(0.1,0.1,0.1,0.1,0.1))

mat<-rdirichlet(10,c(1,1,1))

s<-seq(0,1,by=0.05)
sg<-expand.grid(s,s)
sgm<-sg[rowSums(sg)<=1,]

sdiff <- 1-rowSums(sgm)

mat<-cbind(sgm,sdiff)

z<-ddirichlet(mat,c(2,2,2))

dat<-cbind(x=mat[,1],y=mat[,3],z) %>% as_tibble()

plot_ly(dat, x=~x, y=~y, z= ~z, type="scatter3d", mode="markers", color=~z, size=1)

```

##  example 2a

```{r ex2a.1}
set.seed(762)
n <- 200
x1 <- c(rep(0,n/2), rep(1,n/2))
y <- rnorm(n) + 3*x1
y_tr <- log(y+5)
dat_d<-data.frame(y=ordered(y),x1, y_tr=ordered(y_tr))

fit_orm_d <- orm(y ~ x1, data=dat_d)
```


--> !! Need to get betas from Bayesian CPMs to match beta from orm !! <--

want small value of alpha, try 1/nlevs where nlevs is number of unique levels
how to justify? what sort of prior puts most weight around 0?

http://mc-stan.org/rstanarm/reference/priors.html

- try hs() prior
- try dirchlet(1/n)

```{r ex2a.2, cache=TRUE}
# try stan_polr with different dirichlet specs and look at posterior checks 
fit_polr1_d <- stan_polr(y ~ x1, data = dat_d, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(1), method="logistic",
               adapt_delta = 0.99)

summary(fit_polr1_d, pars=c("x1"), digits=4)

# NOTE: pp_check doesn't look right because y data are factor levels (1:200) while y_rep are actual data points (-2.2: 5.64)
pp_check(fit_polr1_d)  
pp_check(fit_polr1_d, plotfun = "bars", nreps = 50, prob = 0.5)
pp_check(fit_polr1_d, plotfun = "bars_grouped", group = dat2$x1, nreps = 50, prob = 0.5)

fit_polr2_d <- stan_polr(y ~ x1, data = dat_d, iter=3500, 
                       warmup=1500, chains=2,
               prior=NULL, prior_counts = dirichlet(1/200), method="logistic",
               adapt_delta = 0.99)

summary(fit_polr2_d, pars=c("x1"), digits=4)

pp_check(fit_polr2_d)
```


```{r}
# try hardcoded models
mod_data_d <- list(N=nrow(dat_d),
                 ncat=length(unique(dat_d$y)),
                 Ylev=as.numeric(dat_d$y),
                 link=1,
                 K=ncol(dat_d[,c("x1"),drop=FALSE]),
                 Q=dat_d[,c("x1"),drop=FALSE],
                 alpha=1/200)

fit_ord1_d <- sampling(ord_mod1, data=mod_data_d, seed=12345, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_ord2_d <- sampling(ord_mod2, data=mod_data_d, seed=12345, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))
```

```{r}
# compare
fit_orm_d
summary(fit_polr1_d, pars=c("x1"), digits=4) # dirichlet(1),  way off
summary(fit_polr2_d, pars=c("x1"), digits=4) # dirichlet(1/ncat), ok
summary(fit_ord1_d, pars=c("b[1]"))$summary # alpha=1/ncat <-- this one looks best
summary(fit_ord2_d, pars=c("b[1]","alpha"))$summary # est alpha, way off
```


```{r}
fit_orm_e <- orm(y_tr ~ x1, data=dat_d)

mod_data_e <- list(N=nrow(dat_d),
                 ncat=length(unique(dat_d$y_tr)),
                 Ylev=as.numeric(dat_d$y_tr),
                 link=1,
                 K=ncol(dat_d[,c("x1"),drop=FALSE]),
                 Q=dat_d[,c("x1"),drop=FALSE],
                 alpha=1/200)

fit_ord1_e <- sampling(ord_mod1, data=mod_data_d, seed=12345, 
                 iter=3500, warmup=1500, chains=2,
                 control = list(adapt_delta = 0.8))

fit_orm_e
summary(fit_ord1_e, pars=c("b[1]"))$summary



ord1_e_mn_ints <- summary(fit_ord1_e, pars=paste0("cutpoints[",1:199,"]"))$summary[,"mean"]

ord1_e_md_ints <-summary(fit_ord1_e, pars=paste0("cutpoints[",1:199,"]"))$summary[,"50%"]

ord1_e_ci_ints <-summary(fit_ord1_e, pars=paste0("cutpoints[",1:199,"]"))$summary[,c("2.5%", "97.5%")]

# compare intercepts from orm and stan model
ints<-cbind(orm=coef(fit_orm_e)[1:199],
            bayes_mn=-ord1_e_mn_ints,
            bayes_md=-ord1_e_md_ints,
            ci=-ord1_e_ci_ints)

# coefs for rstanarm are negative of coefs from lrm, orm 
qplot(ints[,"orm"], ints[,"bayes_mn"],
      xlim=c(-15,10), ylim=c(-15,10)) + 
  xlab("orm intercept coefs") + ylab("-1 * bayes CPM mean intercept coefs") +
  geom_abline(slope=1,intercept=0)

qplot(ints[,"orm"], ints[,"bayes_md"],
      xlim=c(-15,10), ylim=c(-15,10)) + 
  xlab("orm intercept coefs") + ylab("-1 * bayes CPM median intercept coefs") +
  geom_abline(slope=1,intercept=0)

ints %>% as_tibble() %>% 
  mutate(row=1:nrow(ints)) %>%
  ggplot(aes(x=row,y=orm))+
  geom_point(aes(x=row,y=bayes_md),col=2, pch=1, size=2)+
  geom_point(size=0.5)+
  geom_ribbon(aes(ymin=`97.5%`,ymax=`2.5%`), fill=2, alpha=0.3)

# compare intercepts for both mods
library(tidyr)
ints_plt <- ints %>% as_tibble() %>% 
  mutate(row=1:nrow(ints)) %>% 
  gather(colnames(ints),key="mod",value="value") 

# all are nearly identical except in tails
ints_plt %>% filter(!mod%in%c("2.5%","97.5%")) %>% ggplot(aes(x=row,y=value, color=mod)) + geom_point()

ints_plt %>% filter(!mod%in%c("bayes_mn")) %>% ggplot(aes(x=row,y=value, color=mod)) + geom_point()

```


```{r}
fit_ord1_e

mod_data_e

# Code to get CDF, see function below

#! need to save true y data in mod_data_e 
truey0 <- as.numeric( levels(dat_d$y_tr) ) %>% sort()
  
# prepend value less than min(y) for alpha_0=-Inf intercept
truey<-c(-Inf,truey0) 

# format newdata, betas, and intercepts
newdata=data.frame(`b[1]`=c(0,1))
#! need function to extract beta coef names
cv_nms<-"b[1]" 
ndr <- newdata %>% mutate(ndrow=1:n())
nd <- ndr %>% dplyr::select(-ndrow) %>% as.matrix()

ctpt<-paste0("cutpoints[",1:199,"]")

beta <- as.data.frame(fit_ord1_e) %>% dplyr::select(cv_nms) %>% as.matrix()
int <- as.data.frame(fit_ord1_e) %>% dplyr::select(ctpt) %>% as.matrix()

# get matrix of linear predictions Xb
# (rxp)x (pxs) = rxs
# r is rows in newdata, p is parameters (cols) in newdata, 
# s is number of MCMC samples
Xb <- nd %*% t(beta) 

#use inverse function based on family
inv_func <- plogis

#will have 1 for each row of nd
# check model/doc to make sure values are being calculated correctly
# are cutpoints y<= or y< ??
for (i in 1:nrow(nd)){
      tmpcdf0 <- int - t(Xb[rep(i,ncol(int)),, drop=FALSE]) 
      tmpcdf1 <- cbind(`-Inf`=-Inf, tmpcdf0, `Inf`=Inf) # add alpha_0=-Inf and alpha_n = Inf
      tmpcdf <- tmpcdf1 %>% as.data.frame.table() %>% 
        mutate(cdf=inv_func(Freq), ndrow=i) %>%
        cbind(nd[i,,drop=FALSE])
      assign(paste0("cc",i), tmpcdf)
  }
   
#  F(y_1|X)=G^-1(alpha_i-betaX)

# combine conditional cdfs
nd_ds<-ls()[grep("cc",ls(),fixed=TRUE)] # list of all conditional cdf datasets
cdf_vals<-do.call(rbind, lapply(nd_ds, function(x) get(as.character(x))))

summ<-TRUE
if (summ){
    cdf_summ<-cdf_vals %>%
      ungroup() %>%
      group_by(ndrow, Var2) %>%
      dplyr::summarize(mn_cdf=mean(cdf),
                       med_cdf=median(cdf),
                       cdf_q2.5=quantile(cdf,probs=0.025),
                       cdf_q97.5=quantile(cdf,probs=0.975)) %>%
      ungroup() %>% mutate(yval=rep(truey,nrow(nd))) %>%  
      full_join(., ndr, by="ndrow")
  #  return(cdf_summ)
  } else {
   cdf_out <- cdf_vals %>%
      ungroup() %>%
      dplyr::arrange(ndrow, Var1) %>%
      mutate(yval=rep(truey,nrow(nd)*nsamps  ))
  # return(cdf_out)
  }

cdf_summ

```

```{r}
# getCDF function from rstanarm_ord_functions.R

getCDF <- function(fit, newdata, summ=TRUE,...){
  require(dplyr)
  require(stringr)
  
  #check that cumulative model used
  #check newdata is a data.frame
  # check that names in newdata match coefs from model
  
  if( !identical(sort(names(coef(fit))), 
                sort(names(newdata))) ) stop("newdata vars must match model")
  
  # other checks?
  
  nsamps<-attr(summary(spolrfit),"posterior_sample_size")
  
  # get values of outcome from ordered factor to numeric
  
  #!old don't need real name of outcome, fit object always calls it y
  #fmla<-attr(summary(spolrfit),"formula")
  #outcome<-str_remove(as.character(fmla[2]),"ordered") %>% str_sub(2,-2)
  #truey0 <- as.numeric( levels( unlist(spolrfit[outcome]) ) ) %>% sort()
  
  truey0 <- as.numeric( levels(spolrfit$y) ) %>% sort()
  
  # prepend value less than min(y) for alpha_0=-Inf intercept
  truey<-c(-Inf,truey0) 
  
  # format newdata, betas, and intercepts
  cv_nms<-names(coef(spolrfit)) 
  ndr <- newdata %>% mutate(ndrow=1:n())
  nd <- ndr %>% select(-ndrow) %>% as.matrix()
 
  beta <- as.data.frame(spolrfit) %>% select(cv_nms) %>% as.matrix()
  int <- as.data.frame(spolrfit) %>% select(-cv_nms) %>% as.matrix()
  
  # get matrix of linear predictions Xb
  # (rxp)x (pxs) = rxs
  # r is rows in newdata, p is parameters (cols) in newdata, 
  # s is number of MCMC samples
  Xb <- nd %*% t(beta) 
  
  # add Xb to each intercept (4000xints) 
  #dim(int) => s x (ints-1)
  #dim(Xb) => r x s
  
  #use inverse function based on family
  #! add cauchit
  fam <- spolrfit$family
  if (fam=="probit") {
    inv_func <- pnorm
  } else if (fam == "logistic") {
    inv_func <- plogis
  } else if (fam == "loglog") {
    inv_func <- function(y) exp(-exp(-y))
  } else if (fam == "cloglog") {
    inv_func <- function(y) 1-exp(-exp(y))
  } else if (fam == "cauchit") {
    inv_func <- pcauchy #! not sure if this is right
  }
  
  #will have 1 for each row of nd
  # check model/doc to make sure values are being calculated correctly
  # are cutpoints y<= or y< ??
  for (i in 1:nrow(nd)){
      tmpcdf0 <- int - t(Xb[rep(i,ncol(int)),, drop=FALSE]) 
      tmpcdf1 <- cbind(`-Inf`=-Inf, tmpcdf0, `Inf`=Inf) # add alpha_0=-Inf and alpha_n = Inf
      tmpcdf <- tmpcdf1 %>% as.data.frame.table() %>% 
        mutate(cdf=inv_func(Freq), ndrow=i) %>%
        cbind(nd[i,,drop=FALSE])
      assign(paste0("cc",i), tmpcdf)
  }
   
#  F(y_1|X)=G^-1(alpha_i-betaX)
  
# combine conditional cdfs
nd_ds<-ls()[grep("cc",ls(),fixed=TRUE)] # list of all conditional cdf datasets
cdf_vals<-do.call(rbind, lapply(nd_ds, function(x) get(as.character(x))))

  if (summ){
    cdf_summ<-cdf_vals %>%
      ungroup() %>%
      group_by(ndrow, Var2) %>%
      dplyr::summarize(mn_cdf=mean(cdf),
                       med_cdf=median(cdf),
                       cdf_q2.5=quantile(cdf,probs=0.025),
                       cdf_q97.5=quantile(cdf,probs=0.975)) %>%
      ungroup() %>% mutate(yval=rep(truey,nrow(nd))) %>%  
      full_join(., ndr, by="ndrow")
    return(cdf_summ)
  } else {
   cdf_out <- cdf_vals %>%
      ungroup() %>%
      dplyr::arrange(ndrow, Var1) %>%
      mutate(yval=rep(truey,nrow(nd)*nsamps  ))
   return(cdf_out)
  }

}
```

```{r}
cdf_summ %>% ggplot(aes(group=`b.1.`)) +
  geom_ribbon(aes(x=yval, ymin=cdf_q2.5,ymax=cdf_q97.5),fill="grey30", alpha=0.4)+
  geom_step(aes(x=yval,y=mn_cdf))
```