---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bayesCPM

<!-- badges: start -->
<!-- badges: end -->

This package includes functions to fit a Bayesian Cumulative Probability Model (CPM) using the R interface to Stan.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ntjames/bayes_cpm/pkg")
```

## Example

This is a basic example:

```{r example}
library(bayesCPM)
library(dplyr)

## make example data
set.seed(1567)
n <- 100
x1 <- rnorm(n)
y <- 0.9*x1 + rnorm(n)
dat <- data.frame(y=ordered(y),x1) # outcome must be ordered factor

## sample from Bayes CPM model with probit link
fit <- bayes_cpm(y~x1, data=dat, link="probit", refresh=1000)
```

Using the model fit we can get a summary of the posterior conditional mean,

```{r mean}
## posterior conditional mean when x1=1
fit_mn <- getMean(fit, newdata=data.frame(x1=c(1)))
fit_mn
```

a posterior conditional quantile,

```{r quantile}
## posterior conditional 80th quantile when x1=0.5
fit_q80 <- getQuantile(fit, newdata=data.frame(x1=c(0.5)),q=0.8)
fit_q80
```

or plot the median of the posterior conditional CDF. The true CDF is shown for reference.

```{r cdf, fig.height=5, fig.width=5}
## get posterior conditional CDF when x1=1
fit_cdf <- getCDF(fit, newdata=data.frame(x1=c(1))) 

plot(fit_cdf$yval, fit_cdf$med_cdf, type="s", xlab="y", ylab="Conditional CDF")
curve(pnorm(x,mean=0.9*1),-3,3, add=TRUE, col="red")
legend("topleft",legend=c("Bayes CPM estimate","True CDF"),lty=c(1,1),col=c(1,2))
```

<!--
You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.

You can also embed plots, In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!
-->