---
title: "Bayesian Cumulative Probability Models for Continuous and Mixed Outcomes"
author: "Nathan T. James, Frank E. Harrell Jr., Bryan E. Shepherd, ..."
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: chapter
    number_sections: FALSE
    includes:
      in_header: header.html
    config:
      toc:
        before: <li> &nbsp; Bayesian Cumulative Probability Models for Continuous and Mixed Outcomes </li>
        collapse: section
      download: null
      sharing: no    
  bookdown::pdf_book:
    toc: no
    number_sections: FALSE
    includes:
      in_header: preamble.tex
    keep_tex: no
subparagraph: true
documentclass: article 
bibliography: ["zotero_library.bib"]
linestretch: 1.5
biblio-style: apalike
link-citations: yes
csl: vancouver-brackets-only-year-no-issue.csl
---

```{r setup0, include=FALSE}
#For code block want echo=TRUE if (knitr::is_html_output())
# echo=FALSE otherwise (e.g. echo = FALSE if (knitr::is_latex_output()))
knitr::opts_chunk$set(echo = knitr::is_html_output())
wd<-getwd()

# load packages
libs<-c("knitr", "kableExtra", "magrittr", "ggplot2", "rstan", "plotly", "bayesplot", "ggExtra", "xtable")
invisible(lapply(libs, library, character.only = TRUE))
```

```{r htmlTemp3, echo=FALSE, eval=TRUE}
# code-folding solution from https://stackoverflow.com/questions/45360998/code-folding-in-bookdown

codejs <- readr::read_lines("js/codefolding.js")
collapsejs <- readr::read_lines("js/collapse.js")
transitionjs <- readr::read_lines("js/transition.js")
dropdownjs <- readr::read_lines("js/dropdown.js")

htmlhead <- c(
  paste('
<script>',
paste(transitionjs, collapse = "\n"),
'</script>
<script>',
paste(collapsejs, collapse = "\n"),
'</script>
<script>',
paste(codejs, collapse = "\n"),
'</script>
<script>',
paste(dropdownjs, collapse = "\n"),
'</script>
<style type="text/css">
.code-folding-btn { margin-bottom: 4px; }
.row { display: flex; }
.collapse { display: none; }
.in { display:block }
.pull-right > .dropdown-menu {
    right: 0;
    left: auto;
}
.open > .dropdown-menu {
    display: block;
}
.dropdown-menu {
    position: absolute;
    top: 100%;
    left: 0;
    z-index: 1000;
    display: none;
    float: left;
    min-width: 160px;
    padding: 5px 0;
    margin: 2px 0 0;
    font-size: 14px;
    text-align: left;
    list-style: none;
    background-color: #fff;
    -webkit-background-clip: padding-box;
    background-clip: padding-box;
    border: 1px solid #ccc;
    border: 1px solid rgba(0,0,0,.15);
    border-radius: 4px;
    -webkit-box-shadow: 0 6px 12px rgba(0,0,0,.175);
    box-shadow: 0 6px 12px rgba(0,0,0,.175);
}
</style>
<script>
$(document).ready(function () {
  window.initializeCodeFolding("show" === "hide");
});
</script>
', sep = "\n"),
  paste0('
<script>
document.write(\'<div class="btn-group pull-right" style="position: absolute; top: 20%; right: 2%; z-index: 200"><button type="button" class="btn btn-default btn-xs dropdown-toggle" data-toggle="dropdown" aria-haspopup="true" aria-expanded="true" data-_extension-text-contrast=""><span>Toggle All Code</span> <span class="caret"></span></button><ul class="dropdown-menu" style="min-width: 70px;"><li><a id="rmd-show-all-code" href="#">Show All</a></li><li><a id="rmd-hide-all-code" href="#">Hide All</a></li></ul></div>\')
</script>
')
)

readr::write_lines(htmlhead, path = "header.html")
```

## Abstract

! rework abstract based on updated sims and results !
Ordinal cumulative probability models (CPMs) -- also known as cumulative link models -- such as the proportional odds regression model are typically used for discrete ordered outcomes, but can accommodate both continuous and mixed discrete/continuous outcomes since these are also ordered.  Recent papers by Liu et al. [@liu_modeling_2017] and Tian et al. [@tian_empirical_2019] describe the advantages of ordinal CPMs in this setting using non-parametric maximum likelihood estimation (NPMLE). We formulate a Bayesian CPM for continuous or mixed outcome data. Bayesian CPMs have advantages over frequentist CPMs with regard to interpretation, flexibility, and exact inference (within simulation error) for parameters and functions of parameters<!--when the log-likelihood is not well-approximated by a quadratic function-->. We explore characteristics of the Bayesian CPM through simulations and a case study using HIV biomarker data. In addition, we provide the `bayesCPM` R package which implements Bayesian CPM models using the R interface to Stan. [! Add sentence summarizing findings, e.g. models work with larger sample sizes ...]

<!-- comparison to BNP?? -->

# Introduction